---
title: "Assignment_06_HallVictoria"
author: "Victoria Hall"
date: "10/25/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include = FALSE}
## Set the working directory to the root of your DSC 520 directory
setwd("~/GitHub/dsc520")
```


```{r include = FALSE}
## Load the data/r4ds/heights.csv to
heights_df <- read.csv("data/r4ds/heights.csv")
```


```{r include = FALSE}
## Load the ggplot2 library
library(ggplot2)
```

```{r include = FALSE}
## Fit a linear model using the `age` variable as the predictor and `earn` as the outcome
age_lm <- lm(earn ~ age,data=heights_df)
```

## View the summary of your model using `summary()`
```{r echo= FALSE}
summary(age_lm)
```


```{r echo= FALSE}
## Creating predictions using `predict()`
age_predict_df <- data.frame(earn = predict(age_lm, heights_df), age=heights_df$age)
```
## Plot the predictions against the original data
``` {r echo=FALSE}
ggplot(data = heights_df, aes(y = earn, x = age)) +
  geom_point(color='blue') +
  geom_line(color='red',data = age_predict_df, aes(x= age, y=earn))
```
## Mean
```{r echo=FALSE}
mean_earn <- mean(heights_df$earn)
mean_earn
```

## Corrected Sum of Squares Total
```{r echo= FALSE}
sst <- sum((mean_earn - heights_df$earn)^2)
sst
```

## Corrected Sum of Squares for Model
```{r echo= FALSE}
ssm <- sum((mean_earn - age_predict_df$earn)^2)
ssm
```


``` {r echo= FALSE}
## Residuals
residuals <- heights_df$earn - age_predict_df$earn

```

## Sum of Squares for Error
``` {r echo= FALSE}
sse <- sum(residuals^2)
```

## R Squared
```{r echo=FALSE}
r_squared <- ssm/sst
r_squared
```
## Number of observations
``` {r echo=FALSE}
n <- 1192
n
```
## Number of regression parameters
``` {r echo=FALSE}
p <- 2
p
```
## Corrected Degrees of Freedom for Model (p-1)
``` {r echo=FALSE}
dfm <- p-1
dfm
```
## Degrees of Freedom for Error (n-p)
``` {r echo=FALSE}
dfe <- n-p
dfe
```
## Corrected Degrees of Freedom Total:   DFT = n - 1
``` {r echo=FALSE}
dft <- n-1
dft
```

## Mean of Squares for Model:   MSM = SSM / DFM
``` {r echo=FALSE}
msm <- ssm/dfm
msm
```
## Mean of Squares for Error:   MSE = SSE / DFE
``` {r echo=FALSE}
mse <- sse/dfe
mse
```

## Mean of Squares Total:   MST = SST / DFT
``` {r echo=FALSE}
mst <- sst/dft
mst
```
## F Statistic F = MSM/MSE
``` {r echo=FALSE}
f_score <- msm/mse
f_score
```

## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
``` {r echo=FALSE}
adjusted_r_squared <- 1-(1-r_squared)*(n-1)/(n-p)
adjusted_r_squared
```

## Calculate the p-value from the F distribution
```{r  echo= FALSE}
p_value <- pf(f_score, dfm, dft, lower.tail=F)
p_value
```
