---
title: "Week10_ThoracicSurgery"
author: "Victoria Hall"
date: "11/3/2021"
output: pdf_document
---

``` {r include=FALSE}
library(foreign)
library(mlogit)
library(tidyverse)
thoracicdf<-read.arff("data/ThoraricSurgery.arff")
head(thoracicdf)
```

#In our model, we show a number of high value coefficients yet they have high standard errors and large P values that would indicate these aren't great predictors in our model. You see this is the diagnostic variables (DGN) and a few other indicators as well. There are 4 coefficients that seem to be high and statistically significant. These are PRE9T (Has Dysponea before surgery ), PRE14OC14 (largest size tumor),PRE17T (Has diabetes mellitus), and PRE30T (smoker).A model with these 4 predictors would probably the best choices for the model. If we were to pick the strongest of these choices only, I would pick PRE9T which indicates the patient has Dsyponea before surgery
```{r echo=FALSE}
testglm<-glm(Risk1Yr~ .,data=thoracicdf, family=binomial)
options(scipen=999)
summary(testglm)
```

##I have created an updated model using the 4 significant predictors identified previously. In order to to just look at the largest size tumor as a predictor, I created a new column that is a boolean that indicates if the case has a the largest size tumor (OC14), mark as true, otherwise mark as false. 
```{r echo=FALSE}
#Creating new boolean variable to indicate if they have the largest tumor
thoracicdf<-mutate(thoracicdf, "Largest_Tumor" = ifelse(PRE14 == "OC14", TRUE, FALSE))
head(thoracicdf)
finalglm<-glm(Risk1Yr~PRE9+PRE17+PRE30+Largest_Tumor, data = thoracicdf, family = binomial)
summary(finalglm)
```


##In order to test the accuracy of our model, I predicted the values of the cases and created a confusion matrix that shows the actual values of whether the patient died (T) compared to the predictions of our models. I then found the percent accuracy. Our model was 85% accurate in predicting whether the patient died 1 year after treatment or not. 
```{r echo=FALSE}
prediction<-predict(finalglm,thoracicdf,type="response")
prediction
confmatrix<-table(Actual_Value=thoracicdf$Risk1Yr,Predicted_Value=prediction>0.5)
confmatrix
(confmatrix[[1,1]]+confmatrix[[2,2]])/sum(confmatrix)
```