---
title: "Assignment_07_HallVictoria"
author: "Victoria Hall"
date: "10/25/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
## Set the working directory to the root of your DSC 520 directory
setwd("~/GitHub/dsc520")

## Load the `data/r4ds/heights.csv` to
heights_df <- read.csv("data/r4ds/heights.csv")
heights_df
```

```{r include=FALSE}
# Fit a linear model
earn_lm <-  lm(earn ~ height + sex + ed + age + race, data=heights_df)
```
# View the summary of your model
```{r echo=FALSE}
summary(earn_lm)

predicted_df <- data.frame(
  earn = predict(earn_lm, heights_df),
  ed=heights_df$ed, race=heights_df$race, height=heights_df$height,
  age=heights_df$age, sex=heights_df$sex
  )
```

## Compute deviation (i.e. residuals)
```{r echo=FALSE}
mean_earn <- mean(heights_df$earn)
mean_earn
```

## Corrected Sum of Squares Total
```{r echo=FALSE}
sst <- sum((mean_earn - heights_df$earn)^2)
sst
```

## Corrected Sum of Squares for Model
```{r echo=FALSE}
ssm <- sum((mean_earn - predicted_df$earn)^2)
ssm
```

``` {r echo=FALSE}
## Residuals
residuals <- heights_df$earn - predicted_df$earn
```

## Sum of Squares for Error
``` {r echo=FALSE}
sse <- sum(residuals^2)
sse
```

## R Squared
``` {r echo=FALSE}
r_squared <- ssm/sst
r_squared
```

## Number of observations
```{r echo=FALSE}
n <- 1192
n
```

## Number of regression paramater
```{r echo=FALSE}
p <- 8
p
```

## Corrected Degrees of Freedom for Model
```{r echo=FALSE}
dfm <- p-1
dfm
```

## Degrees of Freedom for Error
```{r echo=FALSE}
dfe <-n-p
dfe
```
## Corrected Degrees of Freedom Total:   DFT = n - 1
```{r echo=FALSE}
dft <- n-1
dft
```

## Mean of Squares for Model:   MSM = SSM / DFM
```{r echo=FALSE}
msm <- ssm/dfm
msm
```

## Mean of Squares for Error:   MSE = SSE / DFE
```{r echo=FALSE}
mse <- sse/dfe
mse
```

## Mean of Squares Total:   MST = SST / DFT
```{r echo=FALSE}
mst <- sst/dft
mst
```

## F Statistic
```{r echo=FALSE}
f_score <- msm/mse
f_score
```

## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
``` {r echo=FALSE}
adjusted_r_squared <- 1 - (1-r_squared)*(n-1)/(n-p)
adjusted_r_squared
```
