---
title: "FinalProjectStep2"
author: "Victoria Hall"
date: "11/7/2021"
output: pdf_document
---

---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/hallt/Documents/GitHub/dsc520/FinalProject")
```

# I have narrowed the focus of my study. I will use a combination of economic data, economic opinion data, and presidential encumbency data to predict the political affliation of the winner of the US presidential election.
1. Economic Indicators - GDP, Personal Income
2. Economic Opinion Data - ANES questions relevant to opinions about the economy
3. US Presidents by date and party


## Libaries needed to import data and clean data
1.Foreign (ANES is a .sav file. I will need to read in the SPSS file and convert it to a dataframe)
2.Tidyverse (mostly for dpylr to manipulate, select, and merge dataframes)
3. Stringr (using to extract just year from dates)

``` {r echo=FALSE}
library(foreign)
library(tidyverse)
library(stringr)
```
```{r include=FALSE}
#Reading in data sets
anesdf<-read.spss("anes_timeseries_cdf.sav", to.data.frame = TRUE)
incomedf<-read.csv("DSPIC96.csv")
GDPdf<-read.csv("GDP.csv")
presdf<-read.csv("US presidents listed - ALL NAMES.csv")
```
### Before I merge frames, I going to go ahead and manipulate each of the datasets just to contain what information I need in the format I need.


1. For the ANES data - I used the codebook to find some questions to use as a part of this analysis. I am going to use square bracket syntax to pull out the variables I want to look at rather than continue along with the full large set. I am also going to just use the years the questions were asked and limit my economic and presidential data to just look at that time frame. Additionally - I will recode responses so we have 3 rather than a continous frame. Further details are inline of the code. VCF9220 - Current retrospective opinion on economy as categorical variable with 1 - better, 2- stayed the same, 3-worse. Asked in 1992, 2000, 2004, 2016.VCF9226 - Opinion on unemployment as categorical variable - 1- Much better, 2- Somewhat better, 3-Same, 4- Somewhat worse, 5-Much worse.VCF9053 - Income kept up with cost of living. 5 point scale - 1. Gone up a lot, 5 - Fall behind a lot. VCF0004 - Year of Study.

2. For my economic data I am using two measures of the state of the economy which are personal income and GDP. I have to change the data format and then find the average per year as the indicators are on a monthly basis rather than yearly. 

3. Presidential - I am going to extract year and party membership of the incumbent as a factor. 
```{r}
#Selecting variables of analysis
anesdf1<-anesdf[,c("VCF0004","VCF9220","VCF9226","VCF9053")]
#Removing years where question wasn't asked
anesdf1<-filter(anesdf1, VCF0004 >=1992)
head(anesdf1)
#Checking formatting of variables
str(anesdf1)
#Rather than moving scales of opinion I'm going to convert responses as factors of 1- same, 2- better, 3-worse. Letting the balance of cases as NA. 
anesdf1$VCF9220<-recode(anesdf1$VCF9220, "1. Gotten better" = 2,"2. Stayed the same" = 1, "3. Gotten worse" = 3)
anesdf1$VCF9220<-as.factor(anesdf1$VCF9220)
str(anesdf1$VCF9220)#Checking it worked correctly
head(anesdf1)
anesdf1$VCF9226<-recode(anesdf1$VCF9226, "1. Much better" = 2, "2. Somewhat better" = 2, "3. Same" = 1, "4. Somewhat worse" = 3, "5. Much worse"=3 )
anesdf1$VCF9226<-as.factor(anesdf1$VCF9226)
anesdf1$VCF9053<-recode(anesdf1$VCF9053, "1. Gone up a lot" = 2, "Gone up a little" = 2, "6. Gone up, DK/NA how much" = 2, "3. Same" = 1, "4. Fallen behind a little" = 3, "5. Fallen behind a lot" = 3, "7. Fallen behind, DK/NA how much" = 3)
anesdf1$VCF9053<-as.factor((anesdf1$VCF9053))
anesdf1<-rename(anesdf1, "DATE"=VCF0004)
str(anesdf1)


#IncomeDataFrameDates are currently formatted as dd/mm/yyyy. In order to later merge frames, converting dates to years only and only 1992 and after.
head(incomedf)
incomedf$DATE<-str_sub(incomedf$DATE, -4, -1)
incomedf$DATE<-as.numeric(incomedf$DATE)
str(incomedf$DATE)
length(incomedf$DATE)
incomedf<-filter(incomedf,DATE >= 1992)
#Because there are multiple responses in a year for the data, I am going to find the average per year
incomedf1<-aggregate(DSPIC96~DATE,incomedf, mean)
head(incomedf1)
#Running the same mutations on the GDP data set
head(GDPdf)
GDPdf$DATE<-str_sub(GDPdf$DATE, -4, -1)
GDPdf$DATE<-as.numeric(GDPdf$DATE)
GDPdf<- filter(GDPdf, DATE>= 1992)
GDPdf<-aggregate(GDP~DATE,GDPdf,mean)
head(GDPdf)

#Presidential Dataframe. Only keeping years after 1992 and refactoring party id.
head(presdf)
presdf<-filter(presdf,Years..after.inauguration.>=1992)
presdf<-rename(presdf, "DATE"=Years..after.inauguration.)
presdf$Party<-recode(presdf$Party,"Democrat"=1, "Republican"=2)
presdf$Party<-as.factor(presdf$Party)
str(presdf)

#Merging 4 df together by date
mergedf<-merge(incomedf1,GDPdf, by="DATE")
mergedf<-merge(mergedf,presdf,by="DATE")
mergedf<-merge(mergedf,anesdf1, by="DATE")
head(mergedf)
```
# Next Steps

### There will be a few different ways to look at this data now that it is merged as one dataframe. I will more than likely need further mutations of the data. One option is look at the change over time in the economy for each party and then find the average increase or decrease of each indicator. I can compare to each value to the mean and possibly create a factor out of this to say if it is good or bad based on it being lower than higher than the mean. 

### It will also be helpful to determine what kind of correlation there is between the actual state of the economy and the opinion data. If there is autocorrelation it might be necessary to run separate models where one uses the actual state of the economy and the other is the opinion data about the economy and perform an anova test on which model performs better. 

### I will probably include a few plots and tables. I would like to plot with color coding the election years and the party the currently is in office. I will also plot some relationships with scatter plots like opinion data as a determiner of the election. It will be very important to distinguish periods and not just use a continous variable of the economic indicators. Things like inflation will mean the the economy will look stronger and stronger each year which is why I want to measure performance in periods based on the prior incumbent party.

### I think will end up using a logistic regression model due to the binomial nature of the outcome (republican or democrat in the US's dominant two party system)


